Handling String-Number Conversion in String Constraint Solving

### 2019-11-06 draft.tex
little things

\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\sti}[1]{\mbox{\textbf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textbf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{\mbox{$V^1_{\Sigma_\epsilon}$}}
\newcommand{\cvartwo}{\mbox{$V^2_{\Sigma_\epsilon}$}}
\newcommand{\modelof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
\newcommand{\parikh}{\mathit{Parikh}}

@@

We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the sets of natural numbers (including 0) and 
integers. For a set $A$, we use $|A|$ to denote its size. 
An \emph{alphabet} is a finite set $\Sigma$ of \emph{characters} and a \emph{string} over $\Sigma$ is a sequence $w = a_1\ldots a_n$ of characters from $\Sigma$, with $\epsilon$ denoting the \emph{empty string}. 
We use $w_1\cdot w_2$ to denote the \emph{concatenation} of strings $w_1$ and $w_2$.
$\Sigma^*$ is the set of all strings over $\Sigma$, $\Sigma^+ = \Sigma^*\setminus \{\epsilon\}$, and 
A \emph{language} over $\Sigma$ is a subset $L$ of $\Sigma^*$.  
%
We use $|w|$ to denote the lenght of $w$ and $|w|_a$ to denote the number of occurrences of the character $a\in \Sigma$ in $w$.  
The \emph{Parikh image} of $w$ is the assignment 
%$\parikh(w) = \{a_1 \mapsto |w|_{a_1},\ldots, a_n \mapsto |w|_{a_n}\}$ that maps every symbol in $w$ to the number of its occurences in $w$. 
$\parikh(w)$ that maps every character $a\in\Sigma$ to the number of its occurences in $w$, i.e., $\parikh(a) = |w|_a$. 
The Parikh image of a language $L$ is the set of Parikh images $\parikh(L) = \{\parikh(w) \mid w \in  L\}$.

@@

The purpose of the structural restrictions of SFA is to ensure that the words of $L(A)$ are uniquelly determined by their Parikh images, as formalised by the following lemma.

\begin{lemma}
%Distict words from the language of an SFA must have distinct Parikh images.  
For an SFA $A$ and $x,y\in L(A)$ if $\parikh(x) = \parikh(y)$ then $x = y$. 
%Distict words from the language of an SFA have distinct Parikh images.  
\end{lemma}
Indeed, notice that each accepted symbolic string $x$ is determined by the numbers of repetitions of individual cycles in the accepting run of $A$ over $x$.
Since every variable appears on at most one transition, 
the number of repetitions of each cyclce equals the value each variable on the edges of the cycle has assigned by $\parikh(x)$ (notice also that $\parikh(x)$ must assign to all variables appearing within the same cycle the same value).

%For such kind of automata, the number of occurrence of each symbol in $\cvar$ uniquely characterize an accepting symbolic string. For example, in the example of Figure~\ref{fig:sfa_def}, if the string $v_0(v^0_1v^1_1v^2_1)^2v_1v_{m-1}$ is accepted, then all other strings $w$ with $|w|_{v_0}=|w|_{v_1}=\ldots=|w|_{v_{m-1}}=1$ and also $|w|_{v^0_1}=|w|_{v^1_1}=|w|_{v^2_1}=2$ are not accepted. Moreover, from $|w|_{v_0}=|w|_{v_1}=\ldots=|w|_{v_{m-1}}=1$ and $|w|_{v^0_1}=|w|_{v^1_1}=|w|_{v^2_1}=2$, we can derive $w=v_0(v^0_1v^1_1v^2_1)^2v_1v_{m-1}$ by exploring the automata structure. \todo{add Presburger constraints to allow more expressive power.}

>>>>

\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\sti}[1]{\mbox{\textbf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textbf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{\mbox{$V^1_{\Sigma_\epsilon}$}}
\newcommand{\cvartwo}{\mbox{$V^2_{\Sigma_\epsilon}$}}
\newcommand{\modelof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}

@@

We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the sets of natural numbers (including 0) and 
integers. For a set $A$, we use $|A|$ to denote its size. 
An \emph{alphabet} is a finite set $\Sigma$ of \emph{characters} and a \emph{string} over $\Sigma$ is a sequence $w = a_1\ldots a_n$ of characters from $\Sigma$, with $\epsilon$ denoting the \emph{empty string}. 
We use $w_1\cdot w_2$ to denote the \emph{concatenation} of strings $w_1$ and $w_2$.
$\Sigma^*$ is the set of all strings over $\Sigma$, $\Sigma^+ = \Sigma^*\setminus \{\epsilon\}$, and 
A \emph{language} over $\Sigma$ is a subset $L$ of $\Sigma^*$.  
%
We use $|w|$ to denote the lenght of $w$ and $|w|_a$ to denote the number of occurrences of the character $a\in \Sigma$ in $w$.  
The \emph{Parikh image} of $w$ is the assignment 
%$\parikh(w) = \{a_1 \mapsto |w|_{a_1},\ldots, a_n \mapsto |w|_{a_n}\}$ that maps every symbol in $w$ to the number of its occurences in $w$. 
$\parikhof w$ that maps every character $a\in\Sigma$ to the number of its occurences in $w$, i.e., $\parikhof w (a) = |w|_a$. 
The Parikh image of a language $L$ is the set of Parikh images $\parikhof L = \{\parikhof w \mid w \in  L\}$.

@@

The purpose of the structural restrictions of SFA is to ensure that the words of $L(A)$ are uniquely determined by their Parikh images, as formalised by the following lemma.

\begin{lemma}
%Distinct words from the language of an SFA must have distinct Parikh images.  
For an SFA $A$ and $x,y\in L(A)$ if $\parikhof x = \parikhof y$ then $x = y$. 
%Distict words from the language of an SFA have distinct Parikh images.  
\end{lemma}
Indeed, notice that each accepted symbolic string $x$ is determined by the numbers of repetitions of individual cycles in the accepting run of $A$ over $x$.
Since every variable appears on at most one transition, then the $\parikhof x$ value of all variables appearing within the same cycle is the same, and it is equal to the number of repetitions of that cycle in the accepting run. 
%
The accepting run on $x$ (and so $x$ itself) can thus be reconstructed from $\parikhof x$.

For example, in the example of Figure~\ref{fig:sfa_def}, 
from $\parikhof x(v_0) = \parikhof x(v_1) = \cdots = \parikhof x(v_{m-1})=1$ and $\parikhof x(v^0_1)=\parikhof x(v^1_1)=\parikhof x(v^2_1)=2$, then $x=v_0(v^0_1v^1_1v^2_1)^2v_1\cdots v_{m-1}$. 
\todo{add Presburger constraints to allow more expressive power.}
